### 2013-03-26 R/reportmd.Rmd
Added automated reports with markdown (first version)

Report for study: `r .study`
========================================================

Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.

Number of records for each variable
-----------------------------------

```{r echo=FALSE}
counts<-apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
counts<-counts[counts>0]  #only include no zeros
counts <- counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
```
```{r echo=FALSE}
cbind(counts)
```

More details
------------------------------------

### Sites

```{r echo=FALSE}
sitevarnames <- var.def$Variable[var.def$Group=="site"]

sitevarnames <- intersect(sitevarnames, names(.dat$data))
```

```{r echo=FALSE, comment=''}
if("location" %in% sitevarnames){
  sites <- sort(levels(.dat$data[,"location"]))
  cat(paste(sites, collapse="\n"))
}
```

```{r echo=FALSE, fig.width=8, fig.height=6}
if("latitude" %in% sitevarnames){
  latlong <- .dat$data[,c("location","latitude","longitude")]
  latlong <- latlong[!duplicated(latlong$location),]

  suppressPackageStartupMessages(library(maps))

  map("world")
  points(latlong$longitude, latlong$latitude, pch=19, col="red")
}
```

### Species

The dataset includes the following species (`r nlevels(.dat$data$species)` species in total):
`r cbind(sort(levels(.dat$data$species)))`

These species belong to the following families:
```{r echo=FALSE}
specvars <- var.def$Variable[var.def$Group=="species"]
specvars <- intersect(specvars, names(.dat$data))
```
```{r echo=FALSE, comment=''}
if("family" %in% specvars){
  families <- sort(unique(.dat$data[,"family"]))
  cat(paste(families, collapse="\n"))
}

```

And include the following plant functional types:
```{r echo=FALSE, comment=''}
if("pft" %in% specvars){
  pfts <- sort(unique(.dat$data[,"pft"]))
  cat(paste(pfts, collapse="\n"))
}

```

Plot of data
---------------------------------
```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
makePlotPanel(d$data, "Ishihara0000", pdf=FALSE)
```