### 2013-03-27 R/reportmd.Rmd
report improved with country maps and more information regarding species traits and what's missing

Report for study: `r .study`
========================================================

Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.

Number of records for each variable
-----------------------------------

```{r echo=FALSE}
counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
counts  <-  counts[counts>0]  #only include no zeros
counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
```
```{r echo=FALSE}
cbind(counts)
```

More details
------------------------------------

### Please analyse if points on the map match locations and coordinates

```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
prepdata  <-  prepMapInfo(alldata$data, study)
drawWorldPlot (prepdata)
```

### Please analyse if points on country map match locations and coordinates

```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
repMissingInfo(prepdata)
drawCountryPlot(prepdata)
```

### Species - please provide any missing information

The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` families, presenting `r length(unique(.dat$data$pft))` functional types, growing in `r length(unique(.dat$data$growingCondition))` conditions within `r length(unique(.dat$data$vegetation))` major types of habitat:
```{r echo=FALSE, message=FALSE, warning=FALSE}
fam.spec(.dat$data)
```

Plot of data
---------------------------------
```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
makePlotPanel(d$data, .study, pdf=FALSE)
```