### 2013-04-09 R/reportmd.Rmd
Improved formatting of knitr report

# Report for study: `r .study`

**Reference:** `r .dat$ref[2]`

**Data contributor:** `r .dat$contact$name`.

**Email:** `r .dat$contact$email`.

**Address:** `r .dat$contact$address`.

## Overview of data provided

Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.

The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` family(ies), presenting `r length(unique(.dat$data$pft))` functional type(s), growing in `r length(unique(.dat$data$growingCondition))` condition(s) within `r length(unique(.dat$data$vegetation))` major type(s) of habitat:

### Number of records for each variable

```{r echo=FALSE, comment=''}
counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
counts  <-  counts[counts>0]  #only include no zeros
counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
cbind(counts)
```

## Study details

### Location

**Location dtaa**
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
locLevelInfo(.dat$data)
```

**Global map**

```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
prepdata  <-  prepMapInfo(alldata$data, study)
drawWorldPlot (prepdata)
```

**Country map**

```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
repMissingInfo(prepdata)
drawCountryPlot(prepdata)
```

**Stand description**

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
standLevelInfo(.dat$data)
```

### Species data

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
spLevelInfo(.dat$data)
```

### Methods
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
metaTable  <-  printMeta(.dat$data)
for(j in 1:nrow(metaTable)){
  cat(paste0(metaTable[j,1], ": ", metaTable[j,2], "\n\n"))
}

```

## Plots of data

This is how the study `r .study` (red symbols) fits in the entire dataset.

```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
makePlotPanel(d$data, .study, pdf=FALSE, quiet=TRUE)
```