### 2013-05-28 talk/slides.tex

\documentclass[t]{beamer}

\mode<presentation>
{
\usetheme{Singapore}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
%\usepackage{graphicx}
\usepackage{movie15}  % For embedding movies http://pages.uoregon.edu/noeckel/PDFmovie.html
\usepackage[normal,tight,center]{subfigure}
\usepackage{epstopdf}
\usepackage{colortbl}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\setbeamertemplate{navigation symbols}{}  % remove navigation material in footer

\title[] % (optional, use only with long paper titles)
{ }
\author%[D~Falster] % (optional, use only with lots of authors)
{Daniel Falster\inst{1}\\
Collaborators: Remko Duursma\inst{2}, Rich FitzJohn\inst{1}, Diego Barneche\inst{1}, many others }
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different affiliation.

\institute[Macquarie University, Sydney] % (optional, but mostly needed)
{
  \inst{1}%
  Biological Sciences,\\
  Macquarie University, Sydney, Australia
  \and
  \inst{2}%
  Hawkesbury Institue for the Environment, \\
  University of Western Sydney, Australia
}

\date[Labchat 2013] % (optional, should be abbreviation of conference name)
{}

\subject{Theoretical Evolutionary Ecology}

%----------------------------------------------------------------------------------------
\begin{document}
%----------------------------------------------------------------------------------------

\titlepage{}

\section{Problem} \subsection{intro}
\begin{frame}[c]  \frametitle{}
\end{frame}

\end{document}