### 2014-07-11 ms/ms.Rmd
few changes/suggestions to ms, changed language to US English

---
output: html_document
---
# *Ecological Archives* Exxxxx
# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx

```{r, results="hide", echo=FALSE, message=FALSE}
source("MS-fun.R")

knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")

bib <- read.bibtex("refs.bib")
baad <- readRDS("../output/baad.rds")
data <- baad$data
var.def <- read.csv("../config/variableDefinitions.csv",
                    stringsAsFactors=FALSE)

vars <- setdiff(var.def$variable[var.def$type == "numeric"], c("map","mat","lai","age"))
```

```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(paste("**List of Authors:**", pasteC(listOfAuthors(baad$contact), ', ')))
```

## INTRODUCTION

Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`. The allocation of reduced carbon into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`. While many scientific studies have quantified plant traits via destructive methods, most of these data have only been made available in the form of summary tables or figures included in publications. Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world. Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale. Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.

We (D.S. Falster and R.A. Duursma, with the help of D.R. Barneche, R.G. FitzJohn and A. V??rhammar) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available. The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data. Moreover, we were invited to incorporate another compilation led by M. Ishihara and focusing on Japanese literature. As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.

Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))`. This compilation would not have been possible without several new technologies. In particular,

1. Our entire work flow is scripted, using the R computing environment. The database can be rebuilt in an instant from the raw data files, using a single command. When a single data contribution has been modified, the final master dataset is thus easily recreated.
2. We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) `r citet("https://github.com/dfalster/dataMashR")` for merging data from different studies in a transparent and reproducible fashion. While it was written with BAAD in mind, the package is designed to be easily  deployed for other compilations.
3. We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data. Datasets as well as processing methods inevitably contain errors. With version control, we are able to trace these errors to the source.
4. We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`. These reports were sent to data contributors for verification, and allowed easy identification of errors in the raw data or processing.

Until recently, scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`. Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies. It is exciting to be able to put these parcels together, to create something larger. We hope others will use this dataset and build on it. See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height8}

# World plot colored by vegetation type
cols <- make.transparent(c(
    "blue","brown","forestgreen","darkorange","dodgerblue2",
    "navajowhite4","lightgoldenrod3","lavender"),0.67)
vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
                 "Temperate rainforest","Tropical rainforest",
                 "Tropical seasonal forest","Woodland","Glasshouse / Common garden")

drawWorldPlot(data[data$vegetation == vegs[1],], sizebyn=TRUE, pchcol=cols[1])
for(i in 2:length(vegs)){
  drawWorldPlot(data[data$vegetation == vegs[i],], sizebyn=TRUE, pchcol=cols[i], add=TRUE)
}
drawWorldPlot(data[data$growingCondition == "GH",], sizebyn=TRUE,
              pchcol=cols[8], add=TRUE)
par(xpd=NA)
legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')

```

**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.

## METHODS

## METADATA CLASS I. DATA SET DESCRIPTORS

### A. DATA SET IDENTITY
BAAD: a Biomass And Allometry Database for woody plants

### B. DATA SET IDENTIFICATION CODE:

### C. DATA SET DESCRIPTORS:
**1. Originators**: The study was initiated by D.S. Falster and R.A. Duursma. D.R. Barneche, R.G. FitzJohn and A. V??rhammar assisted with literature searching, data compilation, and coding. A separate compilation focusing on natural forests in Japan, was collected and/or compiled by M. Ishihara, with assistance from T. Hiura, H. Utsugi and H. Tanouchi. All other authors contributed data to the project.

**2. Abstract**: Understanding how different plants are constructed -- i.e. how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere. Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments. While a variety of models dealing with carbon allocation exist **(Cite references?)**, we still lack a synthetic understanding about the underlying processes. This is partly due to the lack of suitable datasets for validating and parameterizing models. To that end, we present the Biomass And Allometry Database (BAAD) for woody plants. BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species. Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication. Thus BAAD contains individual level data from different studies, transformed into standard units and variable names. The transformations were achieved using a new R package dataMashR.  Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100 m in height; and iii) biomass was estimated directly, i.e. not indirectly via allometric equations, except in very large trees where biomass was first sub-sampled and then scaled up to the entire individual. We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, status, light, experimental treatment, vegetation type) where available. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling.

### D. Key words

plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle

## METADATA CLASS II. RESEARCH ORIGIN DESCRIPTORS

### B. SPECIFIC SUBPROJECT DESCRIPTION

```{r,results="asis",echo=FALSE, message=FALSE}
library(plyr)
library(whisker)
details2B  <-  daply(data, .(studyName), function(x) studyDetails(x))
writeLines(details2B,sep="\n\n")
```

## CLASS III. DATA SET STATUS AND ACCESSIBILITY

### A. Status

**Latest update**: xxxx for all data files.

**Metadata status**: Metadata are complete to last update and are stored
with data.

**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in BAAD, and to minimize outliers.

i) Initial analysis by data contributors. The majority of studies have been published and as such have gone through quality assurance by the data contributor.
ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
iii) All species names were matched against online databases with the Taxonstand and taxize R packages, `r citet(c("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
iv) Additional review by contributors using automated reports. For each study, we generated automated reports using the dataMashR R package (XXXX). In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, metadata provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database). Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.

### B. Accessibility

**1. Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/). The version of the data made available for this publication is available here XXXXX. For scientific analyses, we suggest you use this version.

Development versions of the dataset will continue to be available at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).

If you notice any problem with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).

**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study. For queries about the entire dataset, please contact either Daniel Falster or Remko Duursma. Contact details for all authors are available in the attached table. For problems with the dataset, please submit an issue as described above.

**Copyright restrictions**: The dataset is released under the [Creative Commons CC0 license](http://creativecommons.org/choose/zero/?lang=en). As such, it may be freely used and redistributed. When using the dataset, we kindly request that you cite this article, recognizing the hard work that went into collecting the data and the authors' willingness to make it publicly available.

**Costs**: None.

## CLASS IV. DATA STRUCTURAL DESCRIPTORS

### A. Data Set Files

BAAD contains the following files:

- `dictionary.csv`: a table of variable definitions
- `data.csv`: amalgamated dataset, with columns as defined in `dictionary.csv`
- `references.bib`: bibtex entries containing the primary source for each study
- `references.csv`: summary of bibtex file, giving DOI, URL and citation of  the primary source for each study
- `contacts.csv`: list of contacts and affiliations for each study

### B. Variable definitions for data.csv

```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
x <-daply(var.def, 1, function(x)
  	paste0("* ", x$variable,
    "\n\t- Type: ", x$type,
		"\n\t- Label: ", x$label,
		"\n\t- Description: ", x$description,
		"\n\t- Units: ", gsub(".","/", x$unitsMD, fixed=TRUE)
		))

writeLines(x[var.def$variable])
```

**Number of observations by variable**

```{r results='asis', echo=FALSE, warning=FALSE}
options(digits=9)
kable(summaryTable(data, var.def), "markdown", align="l", padding=10)
```

## ACKNOWLEDGMENTS

D.S. Falster, A. V??rhammer and D.R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A. Duursma. R.G. FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).

## REFERENCES
```{r, echo=FALSE, results='asis'}
bibliography(ordering = c("authors", "year", "title", "journal", "volume", "number", "pages", "url"))
```