### 2014-11-24 reports/report.Rmd

```{r, echo=FALSE, results="hide"}
study_name <- data_study$key
```

# Report for study: `r study_name`

## Contact Information

**Data contributor:** `r data_study$contacts$name`

**Email:** `r paste0(data_study$contacts$email[data_study$contacts$email!=""], collapse = ", ")`

**Address:**

`r paste0("- ", data_study$contacts$address, collapse = "\n")`

## Data source

**Citation:** `r data_study$references$citation`

**DOI:** `r md_link_doi(bib_element(data_study$bibtex, "doi"))`

**Abstract:** `r  bib_element(data_study$bibtex, "abstract")`

## Overview of data provided

The dataset includes records for `r nrow(data_study$data)` individuals from `r length(unique(data_study$data$species))` species belonging to `r length(unique(data_study$data$family))` family(ies), presenting `r length(unique(data_study$data$pft))` functional type(s), growing in `r length(unique(data_study$data$growingCondition))` condition(s) within `r length(unique(data_study$data$vegetation))` major type(s) of habitat, with data included for the following variables:

```{r summary_table, results='asis', echo=FALSE}
# From the ms functions:
tab <- summary_table(data_study$data, baad$dictionary)
kable(tab[names(tab) != "Studies"], "markdown", align="l", padding=10)
```

```{r world_map, echo=FALSE, fig.width=12, fig.height=10}
map_info <- prep_map_info(data_study$data, study_name)
drawWorldPlot(map_info)
```

And locally within the country:

```{r country_map, echo=FALSE, fig.width=12, fig.height=10}
report_missing_info(map_info)
drawCountryPlot(map_info)
```

The sites sampled are:

```{r location_level_info, echo=FALSE, comment='', results='asis'}
kable(location_level_info(data_study$data), "markdown", align="l", padding=2)
```

The growing conditions of sampled plants was:

```{r stand_level_info, echo=FALSE, comment='', results="asis"}
kable(stand_level_info(data_study$data), "markdown", align="l", padding=2)
```

## Species sampled

```{r species_level_info, echo=FALSE, comment='', results="asis"}
kable(species_level_info(data_study$data), "markdown", align="l", padding=2)
```

## Methods used

```{r metadata, echo=FALSE, comment='', results="asis"}
metadata <- data_study$metadata
writeLines(sprintf("**%s**: %s\n", metadata$Topic, metadata$Description))
```

## Plots of data

This is how the study `r study_name` fits in the entire dataset (grey). each colour represents a species. A legend of species names with colours is included at the end for reports with 1 < n < 20 species.

```{r variable_plots, fig.height=12, fig.width=12, echo=FALSE}
report_variable_plots(data_study, baad)
```